---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Lily Botha"
date: "02/05/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

The Denny's dataset has 1643 rows and 6 columns. Each row represents a Denny's restaurant. The columns are the variables: address, city, state, zip, longitude, and latitude. 

```{r dennys_dimensions}
dn <- dennys
nrow(dn)
ncol(dn)
```

### Exercise 2

The La Quinta dataset has 909 rows and 6 columns. Each row represents a La Quinta hotel. The columns are the variables: address, city, state, zip, longitude, and latitude. 

```{r laquinta_dimensions}
lq <- laquinta
nrow(lq)
ncol(lq)
```

### Exercise 3

There are La Quinta hotels outside of the US: in New Zealand, Georgia, Turkey, United Arab Emirates, Colombia, and Ecuador. All of the Denny's are located in the US. 

### Exercise 4

One possibility is to filter by longitude and latitude, which would require finding out the coordinate range of the US. I think a simpler option would be to filter by the state abbreviations, and anything outside of the US's 50 states would be cut. Its also possible that the address of locations outside the US is formatted differently, so we could filter that way, but that seems risky. 

### Exercise 5

There are 0 Denny's locations outside the US. 

```{r dennys_us}
dn %>%
  filter(!(state %in% states$abbreviation)) #filter for states that are not in states$abbreviation
```

### Exercise 6

```{r new_country_var}
dn %>%
  mutate(country = "United States") #add a country variable to Denny's dataset and set all observations to "United States"
```

### Exercise 7

There are La Quinta locations in New Zealand, Georgia, Turkey, United Arab Emirates, Colombia, and Ecuador. 

### Exercise 8
























